Diagnosis_Vis

### 2023-04-14 Data_Import.Rmd
Recodes
Created density plots per year!

view(df)

@@

@@

#Extract Year From Intake Date
#Convert to Date
library("lubridate")
Mothership$Intake_1<-mdy(Mothership$Intake_1,locale = "en_US.UTF-8")
Mothership$TxYear<-as.numeric(format(Mothership$Intake_1,"%Y"))

Mothership<-mutate(Mothership,
                   TxYear = as.factor(TxYear),
                   TxYear = recode(TxYear,
                                   "1582"="2014",
                                   "2021"="2020",
                                   "2022"="2020"))

>>>>

@@

                                      "1"="Treatment Complete",
                                      "2"="Treatment Near Complete",
                                      "3"="Insurance Stopped",
                                      "4"="Nonattendence",
                                      "5"="Inapproparite Language",
                                      "6"="Transfered To Impatient",
                                      "7"="Withdrew for External factors",
                                      "8"="Withdrew (AMA)",
                                      "9"="Other",
                                      "10"="Other"),
                    Referal_1 = as.factor(Referal_1),
                    Referal_1 = recode(Referal_1,
                                      "1"="Inpatient",
                                      "2"="PHP",
                                      "3"="APS",
                                      "4"="Psychatrist",
                                      "5"="PCP",
                                      "6"="Another Physican",
                                      "7"="Therapist",
                                      "17"="Other"
                    ),

@@

#Extract Year From Intake Date
#Convert to Date
library("lubridate")
Mothership$Intake_1<-ymd(Mothership$Intake_1)
Mothership$TxYear<-as.numeric(format(Mothership$Intake_1,"%Y"))
Mothership<-mutate(Mothership,
                   TxYear = as.factor(TxYear),
                   TxYear = recode(TxYear,
                                   "1582"="2014",
                                   "2021"="2020",
                                   "2022"="2020"))

### 2023-04-14 Diagnosis_Vis.Rmd
Created density plots per year!

## Create Diagnosis Variables 
```{r}
Mothership<-Mothership %>%
  mutate(
          # Mood Disorders
                     Mood_D = as.factor(
                      ifelse(
                        MDD_P ==1 | Bipolar1_P == 1 | Bipolar2_P == 1 | Dysthymia_P == 1, 1, 0)),
          # Anxiety Disorders
                    Anxiety_D = as.factor(
                      ifelse(
                        GAD_P ==1 | Panic_P == 1 | PanicAgor_P == 1 | Agoraphobia_P == 1 |
                        Social_P == 1 | Specific_P == 1, 1, 0)),
          # Obsessive Compusive Disorders
                #OCD_P is for current OCD
          # Personality Disorders
                #BPD_P is for current BPD
          # Substance Use Disorders
                    Substance_D = as.factor(
                      ifelse(
                        Alcohol_P == 1 | Drug_P ==1, 1, 0)),
          # Trauma or stresor related disorders
                    Trauma_D = as.factor(
                      ifelse(PTSD_P == 1 | Adjustment_P ==1, 1,0)),
          # Psychotic Disorders
                    Psychotic_D = as.factor(
                      ifelse(
                        Schiz_P == 1 | Schizaff_P == 1,1,0)),
            # Eating Disorders
                    Eating_D = as.factor(
                      ifelse(
                        Anorex_P == 1 | Bulim_P == 1 | Binge_P == 1, 1,0)),
                    Disorder_Type = as.factor(
                      ifelse(Anxiety_D == 1, "Anxiety",
                      ifelse(OCD_P == 1, "OCD",
                      ifelse(BPD_P == 1, "BPD",
                      ifelse(Substance_D == 1, "Substance Use",
                      ifelse(Trauma_D == 1, "Stress and Trauma",
                      ifelse(Mood_D == 1, "Mood",
                      ifelse(Psychotic_D == 1, "Psychotic", "Eating")
                      ))))))))

                      # Comorbid_Anx = as.factor(
                      #   ifelse(Anxiety_D == 1 & Mood_D == 1 |
                      #           Anxiety_D == 1 & OCD_C == 1 |
                      #           Anxiety_D == 1 & BPD_C == 1 |
                      #           Anxiety_D == 1 & Substance_D == 1 |
                      #           Anxiety_D == 1 & Trauma_D == 1 | 
                      #           Anxiety_D == 1 & Eating_D ==1, 1, 0)),
                      # Comorbid_Mood = as.factor(
                      #   ifelse(Mood_D == 1 & Anxiety_D == 1 |
                      #           Mood_D == 1 & BPD_C == 1 |
                      #           Mood_D == 1 & BPD_C == 1 |
                      #           Mood_D == 1 & Substance_D == 1 |
                      #           Mood_D == 1 & Trauma_D == 1 | 
                      #           Mood_D == 1 & Eating_D ==1, 1, 0)),
                      # Comorbid_OCD = as.factor(
                      #   ifelse(OCD_C == 1 & Anxiety_D == 1 |
                      #           OCD_C == 1 & Mood_D == 1 |
                      #           OCD_C == 1 & BPD_C == 1 |
                      #           OCD_C == 1 & Substance_D == 1 |
                      #           OCD_C == 1 & Trauma_D == 1 | 
                      #           OCD_C == 1 & Eating_D ==1, 1, 0)),
                      # Comorbid_BPD = as.factor(
                      #   ifelse(BPD_C == 1 & Anxiety_D == 1 |
                      #           BPD_C == 1 & Mood_D == 1 |
                      #           BPD_C == 1 & OCD_C == 1 |
                      #           BPD_C == 1 & Substance_D == 1 |
                      #           BPD_C == 1 & Trauma_D == 1 | 
                      #           BPD_C == 1 & Eating_D ==1, 1, 0)),

@@

#### Amount of disorders treated
```{r}
#overall psychiatric disorders
ggplot(data = subset(Mothership, !is.na(Disorder_Type)),mapping = aes(x = Disorder_Type))+
  geom_bar()+
  ggtitle("Frequency of Psychiatric Disorder Subtypes") +
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  xlab("Psychiatric Disorder Subtype")+
  theme(panel.grid.major.x = element_line(color = "grey",size = 0.5,linetype = 2))+
  theme(panel.background=NULL)+
  scale_y_continuous(name = "Frequency",limits=c(0,6000))+
  coord_flip()
```

>>>>

## Create Diagnosis Variables 
```{r}
Mothership<-Mothership %>%
  mutate(
          # Mood Disorders
                     Mood_D = as.factor(
                      ifelse(
                        MDD_P ==1 | Bipolar1_P == 1 | Bipolar2_P == 1 | Dysthymia_P == 1, 1, 0)),
          # Anxiety Disorders
                    Anxiety_D = as.factor(
                      ifelse(
                        GAD_P ==1 | Panic_P == 1 | PanicAgor_P == 1 | Agoraphobia_P == 1 |
                        Social_P == 1 | Specific_P == 1, 1, 0)),
          # Obsessive Compusive Disorders
                #OCD_P is for current OCD
          # Personality Disorders
                #BPD_P is for current BPD
          # Substance Use Disorders
                    Substance_D = as.factor(
                      ifelse(
                        Alcohol_P == 1 | Drug_P ==1, 1, 0)),
          # Trauma or stresor related disorders
                    Trauma_D = as.factor(
                      ifelse(PTSD_P == 1 | Adjustment_P ==1, 1,0)),
          # Psychotic Disorders
                    Psychotic_D = as.factor(
                      ifelse(
                        Schiz_P == 1 | Schizaff_P == 1,1,0)),
            # Eating Disorders
                    Eating_D = as.factor(
                      ifelse(
                        Anorex_P == 1 | Bulim_P == 1 | Binge_P == 1, 1,0)),
            # Somatic Disroders
                    Somatic_D = as.factor(
                      ifelse(
                        Somat_P == 1 | UndiffSoma_P == 1| Hypochondriasis_P ==1,
                        1,0)),
                    Disorder_Type = as.factor(
                      ifelse(Anxiety_D == 1, "Anxiety",
                      ifelse(OCD_P == 1, "OCD",
                      ifelse(BPD_P == 1, "BPD",
                      ifelse(Substance_D == 1, "Substance Use",
                      ifelse(Trauma_D == 1, "Stress and Trauma",
                      ifelse(Mood_D == 1, "Mood",
                      ifelse(Psychotic_D == 1, "Psychotic", 
                      ifelse(Eating_D == 1, "Eating", 
                      ifelse(Somatic_D == 1, "Somatic", NA
                      )))))))))),

          #emotional disorders
                  Emotional_dis = as.factor(
                    ifelse(Anxiety_D == 1 | OCD_P == 1 | Trauma_D == 1 |
                           Mood_D == 1, "Emotional Disorder", "Other Disorder"))
                  )
 # Comorbid_Anx = as.factor(
                      #   ifelse(Anxiety_D == 1 & Mood_D == 1 |
                      #           Anxiety_D == 1 & OCD_C == 1 |
                      #           Anxiety_D == 1 & BPD_C == 1 |
                      #           Anxiety_D == 1 & Substance_D == 1 |
                      #           Anxiety_D == 1 & Trauma_D == 1 | 
                      #           Anxiety_D == 1 & Eating_D ==1, 1, 0)),
                      # Comorbid_Mood = as.factor(
                      #   ifelse(Mood_D == 1 & Anxiety_D == 1 |
                      #           Mood_D == 1 & BPD_C == 1 |
                      #           Mood_D == 1 & BPD_C == 1 |
                      #           Mood_D == 1 & Substance_D == 1 |
                      #           Mood_D == 1 & Trauma_D == 1 | 
                      #           Mood_D == 1 & Eating_D ==1, 1, 0)),
                      # Comorbid_OCD = as.factor(
                      #   ifelse(OCD_C == 1 & Anxiety_D == 1 |
                      #           OCD_C == 1 & Mood_D == 1 |
                      #           OCD_C == 1 & BPD_C == 1 |
                      #           OCD_C == 1 & Substance_D == 1 |
                      #           OCD_C == 1 & Trauma_D == 1 | 
                      #           OCD_C == 1 & Eating_D ==1, 1, 0)),
                      # Comorbid_BPD = as.factor(
                      #   ifelse(BPD_C == 1 & Anxiety_D == 1 |
                      #           BPD_C == 1 & Mood_D == 1 |
                      #           BPD_C == 1 & OCD_C == 1 |
                      #           BPD_C == 1 & Substance_D == 1 |
                      #           BPD_C == 1 & Trauma_D == 1 | 
                      #           BPD_C == 1 & Eating_D ==1, 1, 0)),

##### df for plots

# df without NAs in date and disorder type
Mothership_Diag <- Mothership %>%
  filter(!is.na(Disorder_Type))%>%
  filter(!is.na(TxYear))

# df with and without depression
No_depressy <- Mothership_Diag %>%
  filter(Disorder_Type != "Mood")
```

#### Amount of disorders treated
```{r}
#overall psychiatric disorders
ggplot(data = subset(Mothership_Diag, !is.na(Disorder_Type)),mapping = aes(x = Disorder_Type, fill=Disorder_Type))+
  geom_bar()+
  ggtitle("Frequency of Psychiatric Disorder Subtypes") +
  theme(axis.text.x = element_text(angle = 45,hjust = 1))+
  xlab("Psychiatric Disorder Subtype")+
  theme(panel.grid.major.x = element_line(color = "grey",size = 0.5,linetype = 2))+
  theme(panel.background=NULL)+
  scale_y_continuous(name = "Frequency",limits=c(0,3500))+
  coord_flip()
```

#### Number of disorders over time.
```{r}
# Libraries
library(ggplot2)
#install.packages("hrbrthemes")
library(hrbrthemes)
library(dplyr)
library(tidyr)
library(viridis)

# Stacked
ggplot(data=subset(Mothership_Diag, !is.na(Disorder_Type)), 
       mapping= aes(x=TxYear, y = stat(count),
                    group=Disorder_Type, fill=Disorder_Type)) +
    geom_density(adjust=1.5,alpha=.4,position="fill") +
    theme(axis.text.x = element_text(angle = 45,hjust = 1))+
     theme(panel.grid.major.y = element_line(color = "grey",size = 0.5,linetype = 2))+
     xlab("Treatment Year") +
  theme(panel.background=NULL) +
    facet_wrap(~Emotional_dis)+
  labs(
    title = "Type of Disorders Treated at the Hospital Per Year",
    subtitle = "Broken Up By Disorder Types") +
    xlab("Treatment Year") +
    ylab("Percent of Patients With Disorder")

# Broken up by disorder. 
ggplot(data=subset(Mothership, !is.na(Disorder_Type)), 
    mapping= aes(x=TxYear,y = stat(count),
                 group=Disorder_Type, fill=Disorder_Type)) +
  geom_density(adjust=1.5,alpha=.4) +
  facet_wrap(~Disorder_Type)+
  theme_classic()+
    theme(
      panel.grid.major.y = element_line(color = "grey",size = 0.5,linetype = 2),
      axis.text.x = element_text(angle = 45,hjust = 1),
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      axis.ticks.x=element_blank(),
      panel.background=NULL) +
  labs(
    title = "Type of Disorders Treated at the Hospital Per Year",
    subtitle = "Broken Up By Dmotional vs Other Disorders") +
    xlab("Treatment Year") +
    ylab("Number of Patients With Disorder")

@@

```{r}
# without depression
ggplot(data=subset(No_depressy, !is.na(Disorder_Type)), 
    mapping= aes(x=TxYear,y = stat(count),
                 group=Disorder_Type, fill=Disorder_Type)) +
  geom_density(adjust=1.5,alpha=.4) +
  facet_wrap(~Disorder_Type)+
  theme_classic()+
    theme(
      panel.grid.major.y = element_line(color = "grey",size = 0.5,linetype = 2),
      axis.text.x = element_text(angle = 45,hjust = 1),
      legend.position="none",
      panel.spacing = unit(0.1, "lines"),
      axis.ticks.x=element_blank(),
      panel.background=NULL) +
  labs(
    title = "Type of Disorders Treated at the Hospital Per Year",
    subtitle = "Broken Up By Dmotional vs Other Disorders") +
    xlab("Treatment Year") +
    ylab("Number of Patients With Disorder")
```
